---
title: "Chapter 1"
author: "Kevin J. Coyle"
format: gfm
editor: visual
---

```{r}
#| label: setup

library(palmerpenguins)
library(ggthemes)
library(pracma)  # for pracma::nthroot() in 1.4.3 Exercises
library(tidyverse)
```

#### 1.2.5 Exercises

1.  How many rows are in `penguins`? How many columns?

```{r}
nrow(penguins)
```

```{r}
ncol(penguins)
```

2.  What does the `bill_depth_mm` variable in the `penguins` data frame describe? Read the help for [`?penguins`](https://allisonhorst.github.io/palmerpenguins/reference/penguins.html) to find out.

    `bill_depth_mm` is "a number denoting bill depth (millimeters)"

3.  Make a scatterplot of `bill_depth_mm` vs. `bill_length_mm`. That is, make a scatterplot with `bill_depth_mm` on the y-axis and `bill_length_mm` on the x-axis. Describe the relationship between these two variables.

```{r}
  ggplot(data = penguins,
         aes(x = bill_length_mm, y = bill_depth_mm,
             color = species, shape = species)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_color_colorblind()

```

These variables are correlated within species.

4.  What happens if you make a scatterplot of `species` vs. `bill_depth_mm`? What might be a better choice of geom?

```{r}
  ggplot(data = penguins, aes(x = bill_depth_mm, y = species)) +
  geom_point()
```

`geom_boxplot()` would be a better geom.

5.  Why does the following give an error and how would you fix it?

```{r}
#| error: true
ggplot(data = penguins) + 
  geom_point()
```

This causes an error because the aesthetics for `x` and `y` are missing. Instead:

```{r}
  ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point()
```

6.  What does the `na.rm` argument do in [`geom_point()`](https://ggplot2.tidyverse.org/reference/geom_point.html)? What is the default value of the argument? Create a scatterplot where you successfully use this argument set to `TRUE`.

`na.rm` removes missing values without a warning if it is set to `TRUE`. The default value is `FALSE`.

```{r}
  ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(na.rm = TRUE)
```

7.  Add the following caption to the plot you made in the previous exercise: “Data come from the palmerpenguins package.” Hint: Take a look at the documentation for [`labs()`](https://ggplot2.tidyverse.org/reference/labs.html).

```{r}
  ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(na.rm = TRUE) +
  labs(
    caption = "Data come from the palmerpenguins package."
  )
```

8.  Recreate the following visualization. What aesthetic should `bill_depth_mm` be mapped to? And should it be mapped at the global level or at the geom level?

```{r}
  ggplot(data = penguins, aes(x = body_mass_g, y = flipper_length_mm)) +
  geom_point(aes(color = bill_depth_mm), na.rm = TRUE) +
  geom_smooth(method = "lm", na.rm = TRUE)
```

9.  Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.

```{r}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g, color = island)
) +
  geom_point() +
  geom_smooth(se = FALSE)
```

10. Will these two graphs look different? Why/why not?

```{r}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point() +
  geom_smooth()

ggplot() +
  geom_point(
    data = penguins,
    mapping = aes(x = flipper_length_mm, y = body_mass_g)
  ) +
  geom_smooth(
    data = penguins,
    mapping = aes(x = flipper_length_mm, y = body_mass_g)
  )
```

They look the same because the aesthetics specified in the geoms in the second plot are inherited by the geoms in the first plot.

#### 1.4.3 Exercises

1.  Make a bar plot of species of penguins, where you assign species to the y aesthetic. How is this plot different?

```{r}
  ggplot(penguins, aes(y = species)) +
  geom_bar()
```

The bars are horizontal.

2.  How are the following two plots different? Which aesthetic, `color` or `fill`, is more useful for changing the color of bars?

```{r}
ggplot(penguins, aes(x = species)) +
  geom_bar(color = "red")

ggplot(penguins, aes(x = species)) +
  geom_bar(fill = "red")
```

`color` changes only the color of the bars' outline. `fill` is more useful here, since it changes the color of the bar.

3.  What does the `bins` argument in [`geom_histogram()`](https://ggplot2.tidyverse.org/reference/geom_histogram.html) do?

The `bins` argument sets the number of bins in the histogram. It defaults to 30.

4.  Make a histogram of the `carat` variable in the `diamonds`
    dataset that is available when you load the tidyverse package.
    Experiment with different binwidths. What binwidth reveals the most
    interesting patterns?

```{r}
bins <- as.integer(pracma::nthroot(nrow(diamonds), 3) * 2)  # Rice's rule
bin_width <- (max(diamonds$carat) - min(diamonds$carat)) / bins
ggplot(diamonds, aes(x = carat)) +
  geom_histogram(binwidth = bin_width)
```

Rice's rule looks like a good choice for calculating the number of bins, and, correspondingly, an interesting `binwidth`.
