---
title: "Chapter 2"
author: "Kevin J. Coyle"
format: gfm
editor: visual
---

```{r}
#| label: setup

library(tidyverse)
```

#### 2.5 Exercises

1.  Why does this code not work?

```{r}
#| error: true

my_variable <- 10
my_varıable
```

The "i" character is different in the two variable names.

2.  Tweak each of the following R commands so that they run correctly:

```{r}

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() +
  geom_smooth(method = "lm")
```

3.  Press Option + Shift + K / Alt + Shift + K. What happens? How can you get to the same place using the menus?

Pressing this key combination presents a window with Keyboard Shortcut Quick Reference help. Using the menus, Help-\>Keyboard Shortcuts Help presents the same window.

4.  Let’s revisit an exercise from the [Section 1.6](https://r4ds.hadley.nz/data-visualize#sec-ggsave). Run the following lines of code. Which of the two plots is saved as `mpg-plot.png`? Why?

```{r}
my_bar_plot <- ggplot(mpg, aes(x = class)) +
  geom_bar()
my_scatter_plot <- ggplot(mpg, aes(x = cty, y = hwy)) +
  geom_point()
ggsave(filename = "mpg-plot.png", plot = my_bar_plot)
```

The first plot is saved to file as `mpg-plot.png` because the plot is saved in the variable `my_bar_plot`, which is assigned to `plot` in the call to `ggsave()`.
